A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
corpus

In this study, we report on a corpus of \CorpusCorePackages R core and
\CorpusPackages CRAN packages. We run \CorpusAllProgramsRnd programs
extracted from those packages as well as from the user-written data-analyses
shared on the Kaggle platform. In this section we describe this corpus and
methodology used to assemble it.

@@

We have selected the TOP \CorpusPackages packages based on the number of
reverse dependencies, \ie how many other packages depend on
them\footnote{Extracted from package metadata using a builtin function}.
This criterion helps us find packages which should be better written and
have better code coverage.  The resulting set contains \CorpusRCodeRnd lines
of R code and \CorpusNativeCodeRnd lines of native code
(C/C++/Fortran)\footnote{Excluding comments and blank lines using
  \url{github.com/AlDanial/cloc}}. Figure~\ref{fig:corpus} shows these
packages, the size of the dots reflects the project's size in lines of code.
The x-axis indicates code coverage in percents and the y-axis gives the
number of call sites to \eval in log scale. Dotted lines indicate
means. Packages with over \CorpusEvalsPackageTreshold eval call sites are
named.

@@

There are three sources of runnable code in a R package. Next to the package
\emph{tests}, there are also \emph{examples} and \emph{vignettes}. Examples
are R code snippets extracted from package documentation into scripts files.
Vignettes are long-form description of package functionality written using
literate-programming style in \LaTeX\xspace or Rmarkdown with chunks of R
code that can be similarly extracted into R scripts. R provides builtin
functions for tangling the embedded code into files.

The selected packages contain \CorpusPackagePrograms programs with
\CorpusPackageProgramsCodeRnd lines of R code:
%
\begin{compactitem}[$-$]
  \item \CorpusExamplesProgramsRnd examples with
  \summaryrnd{CorpusExamplesCode}\footnote{This notation represents $sum$ ($min$
    / $mean$ / $max$)} lines of code,
  \item \CorpusVignettesProgramsRnd vignettes with
  \summaryrnd{CorpusVignettesCode} lines of code.
\end{compactitem}

For each package we use its runnable code to compute the package code
coverage\footnote{Computed using \url{ https://github.com/r-lib/covr}}. On
average, across the \CorpusPackages packages, it is \CorpusMeanExprCoverage.

@@

To represent user-written code in the corpus, we turned to
Kaggle\footnote{\url{https://www.kaggle.com}}, an online platform for
data-science and machine-learning practitioner. Among others, the website allow
people to share data-science competitions, a data-analysis problem together with
data, for which users try to find the best solution. The solutions, called
\emph{kernels}, are then shared back to the platform in the form of either plain
scripts or as Rmarkdown / Jupyter notebooks written in either Julia, Python, R
or SQLite.

One of the most popular competition is about predicting passenger survival on
Titanic\footnote{\url{https://www.kaggle.com/c/titanic}} with \CorpusKaggle
kernels in R (over 1/4 of all available R kernels) which we used for our corpus.

Unlike CRAN, Kaggle is not a curated repository and therefore there are no
guarantees about the quality of the code. After downloading all of the
\CorpusKaggle kernels and extracting the R code from the various
formats\footnote{We use \c{rmarkdown} package to convert from Jupyer
  notebook to Rmarkdown as well as to convert from Rmarkdown to plain R}, we
found that \CorpusDuplicatedKaggle were whole-file duplicates. From
\CorpusRunnableKaggle kernels, \CorpusFailedKaggle failed to execute. Next to
various runtime exceptions, common problems were parse errors and misspelled
package names.

The final set contains \CorpusFinishedKaggle kernels with
\summaryrnd{CorpusFinishedKaggleCode} lines of R code.

\subsection{Summary}

Together, we have assembled a corpus of \CorpusPackages R packages and
\CorpusFinishedKaggle user-written data-analysis scripts. Together it contains
\CorpusAllCodeRnd lines of source code (both and R and native code). Out of
this, \CorpusAllRunnbaleCode lines R code spread across \CorpusAllPrograms R
scripts is the code that we run during the dynamic analysis.

>>>>

In this study, we report on a corpus of \CorpusCorePackages R core and
\CorpusPackages CRAN packages with \CorpusAllCodeRnd lines of source code
(native + R).  We run \CorpusAllProgramsRnd programs extracted from those
packages as well as from the user-written data-analyses shared on the Kaggle
platform. In this section we describe this corpus and methodology used to
assemble it.

@@

We have selected the TOP \CorpusPackages packages based on their reverse
dependencies, \ie how many other packages depend on them.\footnote{Extracted
  from package metadata using a builtin function.}  This helps us find
packages which should be better written and have better code coverage.  The
resulting set contains \CorpusRCodeRnd lines of R code and
\CorpusNativeCodeRnd lines of native code.\footnote{Excluding comments and
  blank lines using \url{github.com/AlDanial/cloc}.}  For each package we
use its runnable code to compute the package code coverage\footnote{Computed
  using \url{ https://github.com/r-lib/covr}}. On average, across the
\CorpusPackages packages, it is \CorpusMeanExprCoverage.
Figure~\ref{fig:corpus} shows these packages, the size of the dots reflects
the project's size in lines of code.  The x-axis indicates code coverage in
percents and the y-axis gives the number of call sites to \eval in log
scale. Dotted lines indicate means. Packages with over
\CorpusEvalsPackageTreshold eval call sites are named.

@@

There are three sources of runnable code in a R package. Next to the package
\emph{tests},\footnote{We had to exclude most tests because the {testthat}
  harness uses \eval and thus causes the entire test to register as an \eval
  call.} there are also \emph{examples} and \emph{vignettes}. Examples are R
code snippets extracted from package documentation into scripts files.
Vignettes are long-form description of package functionality written using
literate-programming style in \LaTeX\xspace or Rmarkdown with chunks of R
code that can be similarly extracted into R scripts. R provides builtin
functions for tangling the embedded code into files.  The selected packages
contain \CorpusPackagePrograms programs with \CorpusPackageProgramsCodeRnd
lines of R code: \CorpusExamplesProgramsRnd examples with
\summaryrnd{CorpusExamplesCode}\footnote{This notation represents $sum$
  ($min$ / $mean$ / $max$)} lines of code, and \CorpusVignettesProgramsRnd
vignettes.%%% with \summaryrnd{CorpusVignettesCode} lines of code.

@@

To represent end-user code in the corpus, we turned to
Kaggle\footnote{\url{https://www.kaggle.com}}, an online platform for
data-science and machine-learning. The website allow people to share
data-science competitions, a data-analysis problem together with data, for
which users try to find the best solution. The solutions, called {kernels},
are then shared back to the platform in the form of either plain scripts or
as notebooks.  One of the most popular competition is about predicting
passenger survival on
Titanic\footnote{\url{https://www.kaggle.com/c/titanic}} with \CorpusKaggle
kernels in R (over 1/4 of all available R kernels) which we used for our
corpus.

Unlike CRAN, Kaggle is not a curated repository and therefore there are no
guarantees about the quality of the code. After downloading all of the
\CorpusKaggle kernels and extracting the R code from the various
formats,\footnote{We use rmarkdown package to convert from Jupyer notebook
  to Rmarkdown as well as to convert from Rmarkdown to plain R.} we found
that \CorpusDuplicatedKaggle were whole-file duplicates. From
\CorpusRunnableKaggle kernels, \CorpusFailedKaggle failed to execute. Next
to various runtime exceptions, common problems were parse errors and
misspelled package names.  The final set contains \CorpusFinishedKaggle
kernels with \summaryrnd{CorpusFinishedKaggleCode} lines of R code.
